---
title: "Vignette"
author: "Kaylee Frazier"
date: "9/27/2021"
output: 
  github_document:
    toc: true
    html_preview: false
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a vignette to demonstrate how to collect data from an [API](https://en.wikipedia.org/wiki/API). I'll be using the [NASA API](https://api.nasa.gov/index.html). With this API, I'm going to create functions to interact with endpoints and explore the data in this API.

# Required Packages 

To use the functions for interacting with the NASA API, I used the following packages:

- [`tidyverse`](https://www.tidyverse.org/): useful features for data science
- [`jsonlite`](https://cran.r-project.org/web/packages/jsonlite/): API interaction
- [`httr`](https://cran.r-project.org/web/packages/httr/index.html): tools for dealing with URLs and HTTP

```{r echo = TRUE, eval= TRUE, warning=FALSE, message=FALSE}
# Load in the packages.
library(tidyverse)
library(jsonlite)
library(httr)
```


# API Interaction Functions

Below are functions I have created to help interact with the [NASA APIs](https://api.nasa.gov/index.html). These APIs provide insights, images, and meaningful information collected by NASA's various resources. 


## `changeToDF`

This function changes a list element to a data frame.

```{r}
changeToDF <- function(x) {
  do.call(rbind.data.frame, x)
}
```

## `CME.Analysis`

This function takes data from the Coronal Mass Ejection (CME) Analysis endpoint. This function selects the type, longitude, latitude, and speed variables. Then, it filters by the type variable. A user can choose what type they are searching for and the function will return that type's longitude, latitude, and speed.

```{r}
#start the function
CME.Analysis <- function(x) {
#get data from CME Analysis API
dataCME <- GET("https://api.nasa.gov/DONKI/CME?startDate=yyyy-MM-dd&endDate=yyyy-MM-dd&api_key=b7Y6xwkXiwoOc5RRa38biqLuuhcwzC2MoGZ70ecM&feedtype=json&ver=1.0")
#parse the data
parsedCME <- dataCME$content %>% rawToChar() %>% fromJSON()
  #change the list to a data frame
dfCME <- changeToDF(parsedCME$cmeAnalyses)
#select and filter the values you want for the data frame
  dfCME %>% filter(type == x) %>% select(type, longitude, latitude, speed)
}

#example
CME.Analysis("C")
```

# `Speed.Analysis`

This function takes data from the CME API and groups the data by the speed of the coronal mass ejection. I split the categories of speed into low medium and high, and the values for the different categories of speed can be customized to whichever speeds the user wants.

```{r}
#start the function
Speed.Analysis <- function(x, y) {
#get data from CME Analysis API
dataCME <- GET("https://api.nasa.gov/DONKI/CME?startDate=yyyy-MM-dd&endDate=yyyy-MM-dd&api_key=b7Y6xwkXiwoOc5RRa38biqLuuhcwzC2MoGZ70ecM&feedtype=json&ver=1.0")
#parse the data
parsedCME <- dataCME$content %>% rawToChar() %>% fromJSON()
#change the list to a data frame
speed.df <- changeToDF(parsedCME$cmeAnalyses)
#select the variables you want and add a new variable "speedPace"
speed.df %>% mutate(speedPace = if_else(speed >= x, "High", 
    if_else(speed >= y, "Medium", "Low"))) %>% group_by(speedPace) %>% select(speedPace, type, latitude, longitude, halfAngle, speed)
}

#example
Speed.Analysis(500, 400)
```

# `solarFlare`

This function grabs data from the solar flare (FLR) endpoint and returns information about the flareID, the source location, the class type, and the region number. Even though the original data went by the variable names above, this function lets you use whichever names you want for the columns with a being the first column, b being the second column, etc.

```{r}
#start the function
solarFlare <- function(a, b, c, d) {
#grab data from the solar flare API
dataFLR <- GET("https://api.nasa.gov/DONKI/FLR?startDate=2000-01-01&endDate=2020-01-01&api_key=b7Y6xwkXiwoOc5RRa38biqLuuhcwzC2MoGZ70ecM")
#parse the data
parsedFLR <- dataFLR$content %>% rawToChar() %>% fromJSON()
#combine variables in the data
FLR <- cbind(parsedFLR$flrID, parsedFLR$sourceLocation, parsedFLR$activeRegionNum, parsedFLR$classType)
#make the data into a data frame
FLRdf <- as_tibble(FLR)
#change the column names
colnames(FLRdf) <- c(a, b, c, d)
FLRdf %>% select(everything())
}

#example
solarFlare("flrID","sourceLocation", "RegionNum", "classType")
```

# `hazardousAsteroid`

This function takes data from the Asteroids - NeoWs API and returns data about charted asteroids. This function allows you to find whichever summary statistic you want, whether it's the mean, median, standard deiation, etc., for the "absolute_magnitude_h" variable when "is_potentially_hazardous" equals true or false. 

```{r}
#start the function
hazardousAsteroid <- function(stat) {
#grab data from the Asteroids API
dataAsteroid <- GET("https://api.nasa.gov/neo/rest/v1/neo/browse?api_key=b7Y6xwkXiwoOc5RRa38biqLuuhcwzC2MoGZ70ecM") 
#parse the data
parsedAsteroid <- dataAsteroid$content %>% rawToChar() %>% fromJSON()
#grab information from the near_earth_objects data frame
Asteroid <- parsedAsteroid$near_earth_objects
#calculate mean magnitude when the asteroid is and is not hazardous
Magnitude <- aggregate(Asteroid$absolute_magnitude_h, list(Asteroid$is_potentially_hazardous_asteroid), FUN=stat)
#make the output a tibble
Magnitude <- as_tibble(Magnitude)
#create column names
colnames(Magnitude) <- c("is_potentially_hazardous_asteriod", "absolute_magnitude_h")
return(Magnitude)
}

#example
hazardousAsteroid(mean)
```

# `findName`

This function takes information from the Asteroids API. In this function you put in any name of the asteroid you want and it returns the asteriod's name, id, designation, and if it was hazardous.

```{r}
#start the function
findName <- function(x) {
#grab data from the Asteroids API
dataAsteroid <- GET("https://api.nasa.gov/neo/rest/v1/neo/browse?api_key=b7Y6xwkXiwoOc5RRa38biqLuuhcwzC2MoGZ70ecM") 
#parse the data
parsedAsteroid <- dataAsteroid$content %>% rawToChar() %>% fromJSON()
#grab information from the near_earth_objects data frame
Asteroid <- parsedAsteroid$near_earth_objects
#make the output a tibble
Asteroid <- as_tibble(Asteroid)
#select the variables you want outputted and filter by the name of the asteroid
Asteroid %>% filter(name_limited == x) %>% select(name_limited, id, designation, absolute_magnitude_h, is_potentially_hazardous_asteroid)
}

#example
findName("Eros")
```

# `fireballDate` 

This function grabs data from the Fireball API. This function then selects all the data from the "data" data set and allows the user to chose the earliest date it wants to look at to the most recent date. 

```{r}
#start the function
fireballDate <- function(x) {
#grab data from the fireball API
dataFireball <- GET("https://ssd-api.jpl.nasa.gov/fireball.api")
#parse the data
parsedFireball <- dataFireball$content %>% rawToChar() %>% fromJSON()
#grab information from the data data set
Fireball <-  parsedFireball$data
#add column names from the fields data
colnames(Fireball) <- parsedFireball$fields
#make the output into a tibble
Fireball <- as_tibble(Fireball)
#select everything and filter by the date 
Fireball %>% filter(date >= as.Date(x)) %>% select(everything())
}

#example
fireballDate("2020-01-01 12:00:00")
```


# Exploratory Data Analysis

This section takes some of the functions above and grabs data from them, so we can see some trends among the data.

First, I'm going to pull information from the Coronal Mass Ejection endpoint using the `CME.Analysis` function to see the different latitude, longitude, and speed values of each type.

```{r}
CME.Analysis("C")
CME.Analysis("S")
```

I am now going to create a numerical summary of speed for each type of coronal mass ejection

Numerical Summery of speed when type = C
```{r}
CME.Analysis("C") %>% summarise(avgSpeed = mean(speed), medSpeed = median(speed), sdSpeed = sd(speed))
```
Numerical Summary of speed when type = S
```{r}
CME.Analysis("S") %>% summarise(avgSpeed = mean(speed), medSpeed = median(speed), sdSpeed = sd(speed))
```
From these two numerical summeries, we see that when the coronal mass ejecetion type is C, the average speed, median speed, and standard deviation of the speed are all greater than when the type is S.

I am also going to create a boxplot for the type of coronal mass ejection vs. the speed of the coronal mass ejection.
```{r}
g <- ggplot(CME.Analysis("C"), aes(x = type, y = speed))
g + geom_boxplot(fill = "#6fa8dc") +
  labs(title = "Speed When Type = C")
g <- ggplot(CME.Analysis("S"), aes(x = type, y = speed))
g + geom_boxplot(fill = "#ec008a") +
  labs(title = "Speed When Type = S")
```
These two box plots confirm what we saw in the numrical summaries with type C having a greater speed mean, median, and standard deviation. This plot also shows us that when the type is S, the interquartile range is greater than when the type is C.


Using the same CME analysis endpoint, in our `speedAnalysis` function, I created a new variable "speedPace". I am going to customize this variable to where speedPace = "High" when speed >= 500, speedPace = "Medium" when speed >= 400, and speedPace = "Low" for any speeds below those. Using this new variable, I am going to create a barplot of speedPace.
```{r}
speedAnalysis <- Speed.Analysis(500, 400)
g1 <- ggplot(speedAnalysis, aes(x = speedPace, fill = speedPace))
g1 + geom_bar() + labs(x = "Categories of Speed", title = "Speed categories for Coronal Mass Ejections") +
scale_fill_manual(values = c("High" = "#ff8200", "Medium" = "#ec008a", "Low" =  "#6fa8dc")) + theme(legend.position = "none")

```
In this bar plot, we see that most of the coronal mass ejections had a low speed that was below 400. Coronal mass ejections with medium speed were just below low speed, and the coronal mass ejections with a high speed over 500 were the smallest group.





